### 2023-06-05 slides/main.tex
sparse-condensed start

% \input{preamble}

\title{{\bfseries Sparse Condensed-Space Interior-Point Methods with Elastic Relaxations on GPUs: Will it Work?}}

\graphicspath{}

\author[Sungho Shin]{
  Sungho Shin\\
  {\normalfont\footnotesize\url{sshin@anl.gov}}
}
\subtitle{}
\institute[ANL/MCS]{
  Argonne National Laboratory
}

\begin{document}

{
  \usebackgroundtemplate{\includegraphics[width=\paperwidth]{TitleANLBlue}}
  \frame[noframenumbering,plain]{
    \titlepage
    \vspace{-.7in}
  }
}

\begin{frame}{A Naive Approach Doesn't Work.}
  We solve the KKT system within IPM using sparse LU solver in \texttt{CUSOLVERRF}.

  \begin{itemize}
  \item Symbolic factorization needs to be performed every 
  \end{itemize}  
\end{frame}

\begin{frame}{Key Idea}
  Consider an NLP of the following form: 
  \begin{align*}
    \mathop{\text{min}}_{x^L\leq x\leq x^U}\;& f(x)\;\\
    \st\;&c(x) = 0.
  \end{align*}
  The problem is relaxed via {\it elastic relaxation}:
  \begin{align*}
    \mathop{\text{min}}_{x^L\leq x\leq x^U, s}\;& f(x)\;\\
    \st\;& 0\leq c(x) + s\leq \epsilon_{\text{ER}},
  \end{align*}
  where $\epsilon_{\text{ER}}\approx \sqrt{\epsilon_{\text{mach}}}$.

  We now observe that the problem has {\it inequality constriants only}. This allows us to apply the {\it condensation strategy}:
  \begin{align*}
    \begin{bmatrix}
      H + \Sigma_x & &J^\top \\
      & \Sigma_s &I \\
      J & I &\\
    \end{bmatrix}
    \begin{bmatrix}
      p^u \\ p^s \\ p^\lambda
    \end{bmatrix}  = -
    \begin{bmatrix}
      r_1 \\ r_2 \\ r_3 ,
    \end{bmatrix}
    \iff \left(H + \Sigma_x+ J^\top \Sigma_s J \right) p^u = -r_1 + J r_2 - J^\top \Sigma_s r_3.
  \end{align*}
  Of course, $J^\top \Sigma_s J$ can be arbitrarily dense, but our favorite problems will still be sparse.
\end{frame}

\begin{frame}{What do we need for Portabiltiy?}
  \begin{itemize}
  \item All we need is a sparse Cholesky solver (preferrably, written in {\tt KernelAbstractions.jl}).
  \end{itemize}
\end{frame}

\begin{frame}{Schur Complement as ``Bandwidth Thickening''}

\end{frame}

{
  \usebackgroundtemplate{\includegraphics[width=\paperwidth]{TitleANLBlue}}
  \frame[noframenumbering,plain]{
    \titlepage
    \vspace{-.7in}
  }
}

\end{document}