### 2023-07-25 paper/speedup-sol.tex
populated all sections

% Recommended preamble:
% \usetikzlibrary{arrows.meta}
% \usetikzlibrary{backgrounds}
% \usepgfplotslibrary{patchplots}
% \usepgfplotslibrary{fillbetween}
% \pgfplotsset{%
%     layers/standard/.define layer set={%
%         background,axis background,axis grid,axis ticks,axis lines,axis tick labels,pre main,main,axis descriptions,axis foreground%
%     }{
%         grid style={/pgfplots/on layer=axis grid},%
%         tick style={/pgfplots/on layer=axis ticks},%
%         axis line style={/pgfplots/on layer=axis lines},%
%         label style={/pgfplots/on layer=axis descriptions},%
%         legend style={/pgfplots/on layer=axis descriptions},%
%         title style={/pgfplots/on layer=axis descriptions},%
%         colorbar style={/pgfplots/on layer=axis descriptions},%
%         ticklabel style={/pgfplots/on layer=axis tick labels},%
%         axis background@ style={/pgfplots/on layer=axis background},%
%         3d box foreground style={/pgfplots/on layer=axis foreground},%
%     },
% }

\begin{tikzpicture}[/tikz/background rectangle/.style={fill={rgb,1:red,1.0;green,1.0;blue,1.0}, fill opacity={1.0}, draw opacity={1.0}}, show background rectangle]
\begin{axis}[point meta max={nan}, point meta min={nan}, legend cell align={left}, legend columns={1}, title={}, title style={at={{(0.5,1)}}, anchor={south}, font={{\fontsize{14 pt}{18.2 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}, align={center}}, legend style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid, fill={rgb,1:red,1.0;green,1.0;blue,1.0}, fill opacity={1.0}, text opacity={1.0}, font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, text={rgb,1:red,0.0;green,0.0;blue,0.0}, cells={anchor={center}}, at={(0.02, 0.98)}, anchor={north west}}, axis background/.style={fill={rgb,1:red,1.0;green,1.0;blue,1.0}, opacity={1.0}}, anchor={north west}, xshift={1.0mm}, yshift={-1.0mm}, width={150.4mm}, height={99.6mm}, scaled x ticks={false}, xlabel={number of variables}, x tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, x tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, xlabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, at={{(ticklabel cs:0.5)}}, anchor={near ticklabel}, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, xmode={log}, log basis x={10}, xmajorgrids={true}, xmin={1042}, xmax={208624}, xticklabels={{$10^{4}$,$10^{5}$}}, xtick={{10000.0,100000.0}}, xtick align={inside}, xticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, x grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, xticklabel pos={left}, x axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}, scaled y ticks={false}, ylabel={speedup}, y tick style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}}, y tick label style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, opacity={1.0}, rotate={0}}, ylabel style={at={(ticklabel cs:0.5)}, anchor=near ticklabel, at={{(ticklabel cs:0.5)}}, anchor={near ticklabel}, font={{\fontsize{11 pt}{14.3 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, ymode={log}, log basis y={10}, ymajorgrids={true}, ymin={0.042449470320200695}, ymax={37.677607781546705}, yticklabels={{$10^{-1}$,$10^{0}$,$10^{1}$}}, ytick={{0.1,1.0,10.0}}, ytick align={inside}, yticklabel style={font={{\fontsize{8 pt}{10.4 pt}\selectfont}}, color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, rotate={0.0}}, y grid style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={0.1}, line width={0.5}, solid}, yticklabel pos={left}, y axis line style={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, line width={1}, solid}, colorbar={false}]
    \addplot[color={rgb,1:red,0.0;green,0.6056;blue,0.9787}, name path={90991ba5-e570-4efd-b7ef-9814b55ce668}, only marks, draw opacity={1.0}, line width={0}, solid, mark={*}, mark size={3.0 pt}, mark repeat={1}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, fill={rgb,1:red,0.0;green,0.6056;blue,0.9787}, fill opacity={0.0}, line width={0.75}, rotate={0}, solid}]
        table[row sep={\\}]
        {
            \\
            1042.0  0.12132541013470992  \\
            1468.0  0.10314692434543314  \\
            4254.0  0.3591290345722146  \\
            5432.0  0.4318689855256274  \\
            11192.0  0.493180776719653  \\
            17128.0  0.800124080241001  \\
            19008.0  0.9921851426883461  \\
            23238.0  1.0782015249239902  \\
            24540.0  1.2249585614191905  \\
            25086.0  1.0314377489080664  \\
            35270.0  1.8122793796033891  \\
            36696.0  1.9603940524376853  \\
            37872.0  1.8070726670030763  \\
            38814.0  1.908074706072124  \\
            41398.0  1.727056190842467  \\
            42532.0  1.8603003250968289  \\
            76804.0  3.761460351555901  \\
            78748.0  1.5909204206596708  \\
            83572.0  2.421056606260939  \\
            85568.0  2.4138090392087532  \\
            96750.0  1.3842870098320392  \\
            117370.0  2.213283662070483  \\
            179562.0  2.511302487858865  \\
            203374.0  0.5769039469634973  \\
            208624.0  4.274713987006297  \\
        }
        ;
    \addlegendentry {total solution}
    \addplot[color={rgb,1:red,0.8889;green,0.4356;blue,0.2781}, name path={45fed1e9-b750-4c50-97d6-e4d2ebf7d495}, only marks, draw opacity={1.0}, line width={0}, solid, mark={square*}, mark size={3.0 pt}, mark repeat={1}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, fill={rgb,1:red,0.8889;green,0.4356;blue,0.2781}, fill opacity={0.0}, line width={0.75}, rotate={0}, solid}]
        table[row sep={\\}]
        {
            \\
            1042.0  0.1489498744199213  \\
            1468.0  0.1727159137624846  \\
            4254.0  0.4340465934605082  \\
            5432.0  0.5434423830639694  \\
            11192.0  0.8004594515472352  \\
            17128.0  2.082689485354689  \\
            19008.0  2.4770197579261883  \\
            23238.0  3.05899341669706  \\
            24540.0  1.8717332637281556  \\
            25086.0  2.906247958020815  \\
            35270.0  4.496479681105164  \\
            36696.0  3.995367978071199  \\
            37872.0  5.152797200922378  \\
            38814.0  4.421274460394101  \\
            41398.0  4.854859953917821  \\
            42532.0  4.4411634259064  \\
            76804.0  11.566047092387956  \\
            78748.0  5.45272118651195  \\
            83572.0  6.085193656249472  \\
            85568.0  6.910004676366815  \\
            96750.0  6.369970804347526  \\
            117370.0  6.95452612891449  \\
            179562.0  10.976999346486657  \\
            203374.0  18.838803890773352  \\
            208624.0  12.675774859535938  \\
        }
        ;
    \addlegendentry {AD}
    \addplot[color={rgb,1:red,0.2422;green,0.6433;blue,0.3044}, name path={597ac343-eab2-47fd-bfb0-3d1869cc7b86}, only marks, draw opacity={1.0}, line width={0}, solid, mark={diamond*}, mark size={3.0 pt}, mark repeat={1}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, fill={rgb,1:red,0.2422;green,0.6433;blue,0.3044}, fill opacity={0.0}, line width={0.75}, rotate={0}, solid}]
        table[row sep={\\}]
        {
            \\
            1042.0  0.12134780833101436  \\
            1468.0  0.08489894064040139  \\
            4254.0  0.3400556528465654  \\
            5432.0  0.4015885378861249  \\
            11192.0  0.4178468802690407  \\
            17128.0  0.6671440282545749  \\
            19008.0  0.8418469098787906  \\
            23238.0  0.8970213864029066  \\
            24540.0  1.1302597892149948  \\
            25086.0  0.8736098235492576  \\
            35270.0  1.6332401723649006  \\
            36696.0  1.8665104859925663  \\
            37872.0  1.5712590526539294  \\
            38814.0  1.7310996666393024  \\
            41398.0  1.5290569916050272  \\
            42532.0  1.7474748794876105  \\
            76804.0  3.488972057355353  \\
            78748.0  1.395946330507901  \\
            83572.0  2.3246624958297284  \\
            85568.0  2.2224099048005086  \\
            96750.0  1.2509789777402023  \\
            117370.0  2.0490766644040312  \\
            179562.0  2.389075257357081  \\
            203374.0  0.48297264689483094  \\
            208624.0  4.199296385639948  \\
        }
        ;
    \addlegendentry {linear solver}
    \addplot[color={rgb,1:red,0.7644;green,0.4441;blue,0.8243}, name path={0daca70f-5467-4208-80b3-fa1e8e4d7149}, only marks, draw opacity={1.0}, line width={0}, solid, mark={star}, mark size={3.0 pt}, mark repeat={1}, mark options={color={rgb,1:red,0.0;green,0.0;blue,0.0}, draw opacity={1.0}, fill={rgb,1:red,0.7644;green,0.4441;blue,0.8243}, fill opacity={0.0}, line width={0.75}, rotate={0}, solid}]
        table[row sep={\\}]
        {
            \\
            1042.0  0.11053845268205038  \\
            1468.0  0.18528018635787544  \\
            4254.0  0.3971318279284334  \\
            5432.0  0.502547137067705  \\
            11192.0  0.7925028620399515  \\
            17128.0  1.4642569252535946  \\
            19008.0  1.5724747212032781  \\
            23238.0  1.8834530645984655  \\
            24540.0  1.588367729468711  \\
            25086.0  1.6835889681072704  \\
            35270.0  2.4543461608339245  \\
            36696.0  2.2839061860763974  \\
            37872.0  2.5758891325324624  \\
            38814.0  2.548435587364985  \\
            41398.0  2.5197402455776663  \\
            42532.0  2.2091833306911117  \\
            76804.0  4.095029335003687  \\
            78748.0  2.4198733504538077  \\
            83572.0  2.638122640974531  \\
            85568.0  2.8374797565002723  \\
            96750.0  2.5465858998015847  \\
            117370.0  2.5350811531961326  \\
            179562.0  3.059535361656725  \\
            203374.0  3.8895762766124165  \\
            208624.0  3.6340740717834334  \\
        }
        ;
    \addlegendentry {solver internal}
    \addplot[color={rgb,1:red,0.502;green,0.502;blue,0.502}, name path={8f65f3c3-f37c-42b0-ad0a-2ff41a891b22}, draw opacity={1.0}, line width={1}, solid, forget plot]
        table[row sep={\\}]
        {
            \\
            5.204406012731031  1.0  \\
            4.1769648153550915e7  1.0  \\
        }
        ;
\end{axis}
\end{tikzpicture}